Applications of Technology and Large Data in Statistics Education and Statistical Graphics

### 2015-05-25 DefenseSlides/DefenseSlidesCommitteeHour.Rnw
edits to slides for the hour of committee only discussion of chapters 3 and 4

\begin{frame}
    \frametitle{Presentation Outline}  
  Quantifying Loss in Binned Scatterplots
    \begin{itemize}
      \item Scatterplot Adaptations
      \item Binning
      \item Loss Functions
      \item Impact of Binning Specification on Loss
      \item Results and Recommendations
    \end{itemize}    
  Shiny Database Sampler
    \begin{itemize}
      \item Development
      \item Applications
      \item Student User Survey
      \item Results
    \end{itemize}

@@

\begin{center}
\large Loss and Binned Scatterplots\\
\normalsize Chapter 4
\end{center}

@@

\center Heavy Over-plotting

@@

3 Main Choices
  \begin{itemize}
    \item Number of frequency groups (k)
    \item Transform counts (yes/no)
    \item Binning algorithm (standard/quantile)
  \end{itemize}

@@

\begin{frame}
    \frametitle{Number of Frequency Groups}

@@

\begin{frame}
    \frametitle{Frequency Loss Simulation}

@@

\begin{frame}
    \frametitle{Log Frequency Loss Simulation}

@@

<<SkewCounts, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
binout1 <- RectBin2d(datlist[[2]]$x,datlist[[2]]$y,0,0,5,5,type="standard")
p1 <- qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
      main="No Freq Binning") +
  xlab("") + ylab("")+
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43",
                      guide = guide_colourbar(label.position="bottom", label.hjust=0.5,  
                                              title.position = "top")) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     panel.grid= element_blank(),
                     plot.margin = unit(c(.1,.1,.1,.1),"cm") )
p2 <- qplot(binfreq, geom="histogram", data=binout1[[1]], main="Histogram of Raw Bin Counts")+
  theme_bw() + xlab("Bin Frequency") + ylab("Number of Bins")
grid.arrange(p1,p2, nrow=1)
@

\center Removes ability to see differences in low density bins

@@

Transforming bin counts is one way to reduce visual impact of high density bins

@@

\begin{center}
\large Shiny Database Sampler \\
\normalsize Chapter 3
\end{center}

@@

\begin{frame}
    \frametitle{Survey}
<<fluctuationDiagrams,echo=F, message=FALSE, include=T,eval=T,fig.width=12, fig.height=6, out.width='.98\\linewidth', fig.pos='hbt',fig.align='center',tidy=F, cache=TRUE, fig.show='hold'>>=
#ease of use
e12 <- ggfluctuation(table(cdat$q1,cdat$q4)) +
  xlab("Ease Item 1") + ylab("Ease Item 2") + theme_fluct
e13 <- ggfluctuation(table(cdat$q1,cdat$q7)) + xlab("Ease Item 1") + 
  ylab("Ease Item 3") + theme_fluct
e14 <- ggfluctuation(table(cdat$q1,cdat$q10)) + xlab("Ease Item 1") + 
  ylab("Ease Item 4") + theme_fluct
e23 <- ggfluctuation(table(cdat$q4,cdat$q7)) + xlab("Ease Item 2") + 
  ylab("Ease Item 3") + theme_fluct
e24 <- ggfluctuation(table(cdat$q4,cdat$q10)) + xlab("Ease Item 2") + 
  ylab("Ease Item 4") + theme_fluct
e34 <- ggfluctuation(table(cdat$q7,cdat$q10)) + xlab("Ease Item 3") + 
  ylab("Ease Item 4") + theme_fluct

@@

>>>>

\begin{frame}
    \frametitle{Presentation Outline}  
Chapter 4: Binning Strategies and Related Loss for Binned Scatterplots
    \begin{itemize}
      \item Scatterplot Adaptations
      \item Binning and Loss Functions
      \item Impact of Binning Specification on Loss
      \item Results and Recommendations
    \end{itemize}   
\vspace{.2in}    
Chapter 3: A \texttt{shiny} New Opportunity for Interaction with Big Data in Undergraduate Education
    \begin{itemize}
      \item Development
      \item Applications
      \item Student User Survey
      \item Results
    \end{itemize}

@@

\begin{center}
\large Binning Strategies and Related Loss for Binned Scatterplots\\
\normalsize Chapter 4
\end{center}

@@

\center Heavy over-plotting

@@

3 Main Choices
  \begin{itemize}
    \item Number of frequency bins (k)
    \item Transform counts (yes/no)
    \item Binning algorithm (standard/quantile)
  \end{itemize}

@@

\begin{frame}
    \frametitle{Number of Frequency Bins}

@@

@@

\begin{frame}
    \frametitle{Log Frequency Loss}

@@

<<SkewCounts, echo=FALSE, warning=FALSE, message=FALSE, fig.width=11, fig.height=4, out.width='.99\\linewidth', fig.pos='h',fig.align='center',fig.show='hold'>>=
binout1 <- RectBin2d(datlist[[2]]$x,datlist[[2]]$y,0,0,5,5,type="standard")
p1 <- qplot(binxs, binys, geom="tile", fill=binfreq, data=binout1[[1]],
      main="No Freq Binning") +
  xlab("") + ylab("")+
  scale_fill_gradient("Frequency", low="#56B1F7", high="#132B43",
                      guide = guide_colourbar(label.position="bottom", label.hjust=0.5,  
                                              title.position = "top")) + 
  theme_bw() + theme(legend.position="bottom",aspect.ratio=1,
                     panel.grid= element_blank(),
                     plot.margin = unit(c(.1,.1,.1,.1),"cm") )
p2 <- qplot(binfreq, geom="histogram", data=binout1[[1]], main="Histogram of Raw Bin Counts", binwidth=250)+
  theme_bw() + xlab("Bin Frequency") + ylab("Number of Spatial Bins")
grid.arrange(p1,p2, nrow=1)
@

\center Unable to see differences in low frequency bins

@@

\center Transforming bin counts can reduce skew in bin densities

@@

\begin{center}
\large A \texttt{shiny} New Opportunity for Interaction with Big Data in Undergraduate Education \\
\normalsize Chapter 3
\end{center}

@@

\begin{frame}
    \frametitle{Survey Internal Consistency}
<<fluctuationDiagrams,echo=F, message=FALSE, include=T,eval=T,fig.width=12, fig.height=6, out.width='.98\\linewidth', fig.pos='hbt',fig.align='center',tidy=F, cache=TRUE, fig.show='hold'>>=
#ease of use
e12 <- ggfluctuation(table(cdat$q1,cdat$q4)) +
  xlab("Ease Item 1") + ylab("Ease Item 2") + theme_fluct
e13 <- ggfluctuation(table(cdat$q1,cdat$q7)) + xlab("Ease Item 1") + 
  ylab("Ease Item 3") + theme_fluct
e14 <- ggfluctuation(table(cdat$q1,cdat$q10)) + xlab("Ease Item 1") + 
  ylab("Ease Item 4") + theme_fluct
e23 <- ggfluctuation(table(cdat$q4,cdat$q7)) + xlab("Ease Item 2") + 
  ylab("Ease Item 3") + theme_fluct
e24 <- ggfluctuation(table(cdat$q4,cdat$q10)) + xlab("Ease Item 2") + 
  ylab("Ease Item 4") + theme_fluct
e34 <- ggfluctuation(table(cdat$q7,cdat$q10)) + xlab("Ease Item 3") + 
  ylab("Ease Item 4") + theme_fluct

@@

\begin{frame}
    \frametitle{Internal Consistency}
